Incentivizing Exploration with Unbiased History

### 2018-10-28 3level.tex
check in often!

\section{Adaptive Policy with Three Levels}
\label{sec:3level}
\swcomment{rewrote below}
Now we demonstrate how to design a 3-level policy to incentivize
adaptive exploration and improve the $O(T^{2/3})$ regret rate to
$\tilde O(T^{4/7})$. In \Cref{sec:llevel}, we will show how to extend
this idea to achieve a regret rate of $\tilde O(T^{1/2})$. For
simplicity, we will focus on the setting with 2 arms.

The policy consists of three phases, each of which corresponds to a
level (see Figure~\ref{fig:3level}).  In the first level, the policy
performs the first round of exploration over the first
$(S\cdot T_1\cdot T_G)$ agents, where $T_1 = T^{4/7}\log^{-1/7}(T)$
and $S = 2^{10}\log(T)$. In particular, the policy randomly partitions
the agents into $S$ groups such that each group consists of $T_1$ runs
of \ALGG with $T_G$ rounds. In the second level, the policy performs
further exploration on the next $(S\cdot T_2)$ agents by incorporating
data collected in the first level, where
$T_2 = T^{6/7}\log^{-5/7}(T)$. The agents are again partitioned into
$S$ groups such that the agents in the $s$-th group are shown the same
sub-history from the $s$-th group in the first level. In the third
level, the policy exploits by showing all remaining agents the entire
history from the first two levels.

They main idea for the 3-level policy is that it inserts the second
level as an additional ``checkpoint'' to induce more adaptivity in the
exploration. While the first level ensures that the agents explore
both arms, the second level will continue to explore both arms only if
the rewards of the two arms are still indistinguishable (that is, when
$|\mu_1 - \mu_2| < \tilde{O}(1/\sqrt{T_1})$). In the third level,
agents will pull the best arm unless $|\mu_1-\mu_2|$ is much smaller
than $\tilde{O}(1/\sqrt{T_2})$, in which case pulling either arm will
incur small regret. The formal guarantee of the policy is the
following.

@@

\subsection{Clean events}
We will first establish four \emph{clean events} that together occur
with probability at least $1 - O(1/T)$. The first event says that
every group of \ALGG runs in the first level ensures that each arm $a$
is pulled at least $\Omega(T_1)$ times. The following lemma can
derived from combining Lemma~\ref{lem:t1runs} and union bound.

@@

The next two events are about the distribution of the mean rewards in
the first level. To state the events, it will be useful to think of a
hypothetical reward tape tape $\cT^1_{s, a}$ of length $T$ for each
group $s$ and arm $a$, with each cell independently sampled from
$\cD_a$.  The tape encodes rewards as follows: the $j$-th time arm $a$
is chosen by the group $s$ in the first level, its reward is taken
from the $j$-th cell in this arm’s tape. The following result
characterizes the concentration of the mean rewards among all
consecutive pulls among all such tapes, which follows from Chernoff
bound and union bound.

@@

Our policy also relies on the anti-concentration of the empirical
means in the first round. We show that for each arm $a\in \{1, 2\}$,
there exists a group $s_a$ such that the empirical mean of $a$ is
slightly above $\mu_a$, while the other arm $(3 - a)$ has empirical
mean slightly below $\mu_{(3-a)}$. This event is crucial for making
the agents in the second level to explore both arms when the their
mean rewards are indistinguishable after the first level.

\begin{lemma}[Co-occurence of high and low deviations]\label{3levelw4}
  For any group $s\in [S]$, any arm $a$, let $\tilde\mu_{a,s}$ be the
  empirical mean of the first $q_a T_1$ cells in tape $\cT^1_{a, s}$.
  Let $W_4^{s,a,high}$ be the event
  $\tilde\mu_{a, s} \geq \mu_a + 1/\sqrt{q_a T_1}$ and let
  $W_4^{s,a,low}$ be the event that
  $\tilde\mu_{a, s} \leq \mu_a - 1/\sqrt{q_a T_1}$.  Let $W_4$ be the
  event that for every $a\in \{1, 2\}$, there exists a group
  $s_a\in [S]$ in the first level such that both $W_4^{s_a,a,high}$
  and $W_4^{s_a,3-a,low}$ occur. Then
  \[
    \Pr[W_4]\geq 1 -2 /T.
  \]
\end{lemma}

Lastly, we will condition on the event that the empirical means of
both arms to be concentrated around their true means in any prefix of
their pulls. This guarantees that the policy obtains an accurate
estimate of rewards for both arms after aggregating all the data in
the first two levels.
% We will again state the event using a hypothetical reward tape
% $\cT^2_a$ for each arm $a$: the $j$-th time arm $a$ is chosen in the
% two levels, its reward is taken from the $j$-th cell in $\cT^2_a$.

\swcomment{might want to swap out $W_3$ and $W_4$}
\begin{lemma}[Concentration of empirical means in the first two
  levels]\label{3levelw3}
  With probability at least $1 - \frac{4}{T^3}$, the following event
  $W_3$ holds: for all $a\in \{1, 2\}$ and $m \in [N_{T, a}]$, the
  empirical means of the first $t$ arm $a$ pulls is at most
  $\sqrt{\frac{2\log(T)}{m}}$ away from $\mu_a$, where $N_{T, a}$ is
  the total number of arm $a$ pulls by the end of $T$ rounds.
\end{lemma}

Let $W = \bigcap_{i=1}^4 W_i$ be the intersection of all 4 clean
events.  By union bound, $W$ occuers with probability $1-O(1/T)$. Note
that expected regret conditioned on $W$ not occuring is at most
$O(1/T) \cdot T = O(1)$, so it suffices to bound the regret under $W$.

\subsection{Case Analysis under Clean Events}

We now proceed to bound the regret conditioned on all four clean
events defined above occur. Without loss of generality, we assume
$\mu_1 \geq \mu_2$ as the recommendation policy is symmetric to both
arms. We break down the regret analysis into four cases on the scale
of the gap parameter $\Delta = (\mu_1-\mu_2)$.

@@

Another simple case is when $\Delta$ is sufficiently large, so that
the data collected in any sing first-level group can distinguish the
two arms. The proof essentially follows from the clean event $W_1$
that both arms are explored sufficient many times in level 1, and the
empirical rewards in the first two levels concentrate around their
true means.

@@

Finally, consider the \emph{small gap} case where $\Delta$ is between
$\tilde\Omega(\sqrt{1/T_2})$ and $\tilde O(\sqrt{1/(S T_1))})$. This
is more challenging since even aggregating the data from all $S$
groups in the first level is not sufficient for correctly
distinguishing the two arms. Thus, we need to ensure that both arms
will continue to be pulled in the second level. To achieve this, we
will nicely leverage the event of $W_4$, which has the implication
that every arm $a$ has its ``lucky'' group $s_a$ in the first level,
for which the empirical mean is slightly higher than $\mu_a$, while
the other arm is slightly lower than $\mu_{(3-a)}$. Since the
deviations are in the order of $\Omega(\sqrt{1/T_1})$, the sub-history
from this ``lucky'' group will induce a pair of estimates such that
$\hat \mu_a^t > \hat \mu_{(3-a)}^t$, and so any second-level agent in
group $s_a$ will pull arm $a$.  As a result, we guarantee both arms
will be pulled at least $T_2$ times in the second level, which in turn
gives the following guarantee.

@@

\paragraph{Wrapping up: proof of \Cref{thm:3level}. } In negligible
gap case, the stated regret bound holds regardless of the sequence of
chosen arms. In the large gap case, the regret only comes from the
first level, which is bounded
by$S T_G T_1 = O(T^{4/7} \log^{6/7}(T))$. In the intermediate cases of
both medium and small gap, it suffices to bound the regret from the
first and second levels, so
\[
\reg(T) \leq (S T_G T_1 + S T_2) \cdot 2\left(\sqrt{\frac{4\log(T)}{q_1T_1}} 
+ \sqrt{\frac{4\log(T)}{q_2T_1}}\right) = O(T^{4/7} \log^{6/7}(T)).
\]
Therefore, we obtain the stated regret bound in all cases.

>>>>

\section{Adaptive Policy with Three Levels}
\label{sec:3level}
\swcomment{rewrote below}
Now we demonstrate how to design a 3-level policy to incentivize
adaptive exploration and improve the $O(T^{2/3})$ regret rate to
$\tilde O(T^{4/7})$. In \Cref{sec:llevel}, we will show how to extend
this idea to achieve a regret rate of $\tilde O(T^{1/2})$. For
simplicity, we will focus on the setting with 2 arms\niedit{, although our results can be extended to a constant number of arms $K$}. \nicomment{I think that's true, but if not, then we're not focusing on 2 arms for ``simplicity''.}

The policy consists of three phases, each of which corresponds to a
level (see Figure~\ref{fig:3level}).  In the first level, the policy
performs the first round of exploration over the first
$(S\cdot T_1\cdot T_G)$ agents, where $T_1 = T^{4/7}\log^{-1/7}(T)$
and $S = 2^{10}\log(T)$. In particular, the policy randomly partitions
the agents into $S$ groups such that each group consists of $T_1$ runs
of \ALGG with $T_G$ rounds. In the second level, the policy performs
further exploration on the next $(S\cdot T_2)$ agents by incorporating
data collected in the first level, where
$T_2 = T^{6/7}\log^{-5/7}(T)$. The agents are again partitioned into
$S$ groups such that the agents in the $s$-th group are shown the same
sub-history from the $s$-th group in the first level. In the third
level, the policy exploits by showing all remaining agents the entire
history from the first two levels. \nicomment{This comment applies to the 2level too; why not allow the guys within the last level see each other?  Seems like this would still work but be more of a headache to prove -- would have to argue the empirical means can't move around too much in the sub-terminal levels, but it makes a big qualitative difference because now everyone sees a vanishingly small fraction of the history whereas with this modification a fraction approaching 1 of the people see a fraction approaching 1 of the history.}

They main idea for the 3-level policy is that it inserts the second
level as an additional ``checkpoint'' to induce more adaptivity in the
exploration. While the first level ensures that the agents explore
both arms, the second level will continue to explore both arms only if
the rewards of the two arms are still indistinguishable (that is, when
$|\mu_1 - \mu_2| < \tilde{O}(1/\sqrt{T_1})$). In the third level,
agents will pull the best arm unless $|\mu_1-\mu_2|$ is much smaller
than $\tilde{O}(1/\sqrt{T_2})$, in which case pulling either arm will
incur small regret. \nicomment{Need to add intuition about why it's necessary to have a middle level to get this to happen, as opposed to making the bottom level bigger.  Is it because the bottom level will start to herd on some arm and we want to refresh that process? Also should talk about need for anti-concentration bound.} The formal guarantee of the policy is the
following.

@@

\subsection{Events}
We will first establish four \nidelete{\emph{clean events}} \niedit{events} \nicomment{this is a dumb name, what's clean or dirty about an event?  i advocate deleting this adjective from our paper, as i did for the 2level section} that together occur
with probability at least $1 - O(1/T)$. The first event says that
every group of \ALGG runs in the first level ensures that each arm $a$
is pulled at least $\Omega(T_1)$ times. The following lemma can
derived from combining Lemma~\ref{lem:t1runs} and union bound.

@@

The next two events are about the distribution of the mean rewards in
the first level. To state the events, it will be useful to think of a
hypothetical reward tape $\cT^1_{s, a}$ of length $T$ for each
group $s$ and arm $a$, with each cell independently sampled from
$\cD_a$.  The tape encodes rewards as follows: the $j$-th time arm $a$
is chosen by the group $s$ in the first level, its reward is taken
from the $j$-th cell in this arm's tape. The following result
characterizes the concentration of the mean rewards among all
consecutive pulls among all such tapes, which follows from Chernoff
bound and union bound.

@@

Our policy also relies on the anti-concentration of the empirical
means in the first round. We show that for each arm $a\in \{1, 2\}$,
there exists a group $s_a$ such that the empirical mean of $a$ is
slightly above $\mu_a$, while the other arm $(3 - a)$ has empirical
mean slightly below $\mu_{(3-a)}$. This event is crucial for inducing
agents in the second level to explore both arms when the their
mean rewards are indistinguishable after the first level.

\begin{lemma}[Co-occurence of high and low deviations]\label{3levelw4}
  For any group $s\in [S]$, any arm $a$, let $\tilde\mu_{a,s}$ be the
  empirical mean of the first $q_a T_1$ cells in tape $\cT^1_{a, s}$.
  Let $W_4^{s,a,\text{high}}$ be the event
  $\tilde\mu_{a, s} \geq \mu_a + 1/\sqrt{q_a T_1}$ and let
  $W_4^{s,a,\text{low}}$ be the event that
  $\tilde\mu_{a, s} \leq \mu_a - 1/\sqrt{q_a T_1}$.  Let $W_4$ be the
  event that for every $a\in \{1, 2\}$, there exists a group
  $s_a\in [S]$ in the first level such that both $W_4^{s_a,a,\text{high}}$
  and $W_4^{s_a,3-a,\text{low}}$ occur. Then
  \[
    \Pr[W_4]\geq 1 -2 /T.
  \]
\end{lemma}

Lastly, we will condition on the event that the empirical means of
both arms are concentrated around their true means in any prefix of
their pulls. This guarantees that the policy obtains an accurate
estimate of rewards for both arms after aggregating all the data in
the first two levels.
% We will again state the event using a hypothetical reward tape
% $\cT^2_a$ for each arm $a$: the $j$-th time arm $a$ is chosen in the
% two levels, its reward is taken from the $j$-th cell in $\cT^2_a$.

\swcomment{might want to swap out $W_3$ and $W_4$} \nicomment{yes please!}

\begin{lemma}[Concentration of empirical means in the first two
  levels]\label{3levelw3}
  With probability at least $1 - \frac{4}{T^3}$, the following event
  $W_3$ holds: for all $a\in \{1, 2\}$ and $m \in [N_{T, a}]$, the
  empirical means of the first $t$ \nicomment{what is $t$ here; you mean $m$? in which case please replace $m$ with $t$ in the conditioning and bound} arm $a$ pulls is at most
  $\sqrt{\frac{2\log(T)}{m}}$ away from $\mu_a$, where $N_{T, a}$ is
  the total number of arm $a$ pulls by the end of $T$ rounds.
\end{lemma}

Let $W = \bigcap_{i=1}^4 W_i$ be the intersection of all 4 
events.  By union bound, $W$ occurs with probability $1-O(1/T)$. Note
that expected regret conditioned on $W$ not occuring is at most
$O(1/T) \cdot T = O(1)$, so it suffices to bound the regret conditioned on $W$.

\subsection{Case Analysis}

We now proceed to bound the regret conditioned on the four 
events defined above. Without loss of generality, we assume
$\mu_1 \geq \mu_2$ as the recommendation policy is symmetric to both
arms. We break down the regret analysis into four cases on the scale
of the gap parameter $\Delta = (\mu_1-\mu_2)$.

@@

Another simple case is when $\Delta$ is sufficiently large, so that
the data collected in any single first-level group can distinguish the
two arms. The proof essentially follows from the event $W_1$
that both arms are explored sufficient many times in level 1, and the
empirical rewards in the first two levels concentrate around their
true means.

@@

Finally, consider the \emph{small gap} case where $\Delta$ is between
$\tilde\Omega(\sqrt{1/T_2})$ and $\tilde O(\sqrt{1/(S T_1))})$. This
is more challenging since even aggregating the data from all $S$
groups in the first level is not sufficient for correctly
distinguishing the two arms. Thus, we need to ensure that both arms
will continue to be pulled in the second level. To achieve this, we
will nicely leverage the event of $W_4$, which has the implication
that each arm $a$ has a ``lucky'' group $s_a$ in the first level,
for which its empirical mean is slightly higher than $\mu_a$, while
the other arm is slightly lower than $\mu_{(3-a)}$. Since the
deviations are in the order of $\Omega(\sqrt{1/T_1})$, the sub-history
from this ``lucky'' group will induce a pair of estimates such that
$\hat \mu_a^t > \hat \mu_{(3-a)}^t$, and so any second-level agent in
group $s_a$ will pull arm $a$.  As a result, we guarantee both arms
will be pulled at least $T_2$ times in the second level, which in turn
gives the following guarantee.

@@

\paragraph{Wrapping up: proof of \Cref{thm:3level}. } In negligible
gap case, the stated regret bound holds regardless of the sequence of
chosen arms. In the large gap case, the regret only comes from the
first level, which is bounded
by $S T_G T_1 = O(T^{4/7} \log^{6/7}(T))$. In the intermediate cases of
both medium and small gaps, it suffices to bound the regret from the
first and second levels, so
\[
\reg(T) \leq (S T_G T_1 + S T_2) \cdot 2\left(\sqrt{\frac{4\log(T)}{q_1T_1}} 
+ \sqrt{\frac{4\log(T)}{q_2T_1}}\right) = O(T^{4/7} \log^{6/7}(T)).
\]
Therefore, we obtain the stated regret bound in all cases.
